2015 NYC Street Tree Census EDA by Sayed Malawi
========================================================
```{r global_options, include=FALSE}
knitr::opts_chunk$set(echo=FALSE, warning=FALSE, message=FALSE)
```

```{r echo=FALSE, message=FALSE, warning=FALSE, packages}
# Load all of the packages that you end up using in your analysis in this code
# chunk.

# Notice that the parameter "echo" was set to FALSE for this code chunk. This
# prevents the code from displaying in the knitted HTML output. You should set
# echo=FALSE for all code chunks in your file, unless it makes sense for your
# report to show the code that generated a particular plot.

# The other parameters for "message" and "warning" should also be set to FALSE
# for other code chunks once you have verified that each plot comes out as you
# want it to. This will clean up the flow of your report.

# install.packages('devtools')
# devtools::install_github("zachcp/nycmaps")
# install.packages('mapproj')
# install.packages('rgdal')
# install.packages('rgeos')
# install.packages('maptools')
# install.packages('sf')
library(ggplot2)
library(dplyr)
library(mapproj)
library(nycmaps)
library(rgdal)
library(maptools)
library(rgeos)
library(sf)
```

```{r echo=FALSE, Load_the_Data}
# Load the Data
trees <- read.csv('Data/2015_Street_Tree_Census_-_Tree_Data.csv')
```

This report contains an exploration of data from the 2015 New York City
TreesCount street tree census, organized by the city's Department of Parks &
Recreation. Data on over 680,000 of the city's street trees (i.e. not including
parks, etc.) was collected by staff and
volunteers. Each observation contains up to 40 pieces of information, such as
species, health, and size, as well as highly accurate location data. As a
student worker at my college's arboretum, this dataset was of interest to me;
additionally, I was inspired by the use of map plots in some of the example
projects.

# Univariate Plots Section

```{r echo=FALSE, Univariate_Plots}
str(trees)
```

There are a lot of factor variables to work with! Many are related to location
and plant bed problems.

We can start with some simple plots investigating one variable from the dataset.

```{r echo=FALSE}
ggplot(aes(x = status), data = trees) +
  geom_bar()
```

Most of the street trees happen to be alive, thankfully. ("Dead" refers to
still-standing dead trees, as opposed to stumps.)

```{r echo=FALSE}
ggplot(aes(x = health), data = subset(trees, status == 'Alive')) +
  geom_bar()
```

Of the trees that are alive, it looks like about 70-80% are in good health.

```{r echo=FALSE}
trees$tree_dbh[trees$status == 'Stump'] <- NA

ggplot(aes(x = tree_dbh), data = subset(trees, !is.na(tree_dbh))) +
  geom_histogram(binwidth = 5) +
  scale_y_log10()
```

This plot shows the diameters of all non-stump trees, measured in inches, with a
log scale applied to the y axis. Evidently, the vast majority of street trees
are on the small or medium side, with diameters in the 0-50 inch range.

```{r echo=FALSE}
summary(subset(trees$tree_dbh, !is.na(trees$tree_dbh)))
```

Some of the street trees with diameters in the hundred-inch range may be victims
of data misentry. Plotting without the outliers may give a better visualization
of the distribution.

```{r echo=FALSE}
ggplot(aes(x = tree_dbh), data = subset(trees, !is.na(tree_dbh))) +
  geom_histogram(binwidth = 2) +
  scale_y_log10() +
  coord_cartesian(xlim = c(0, quantile(trees$tree_dbh, 0.995, na.rm = T)))
```

As the above graph shows, the distribution is still right-skewed, but not as
severely as previously thought.

```{r echo=FALSE}
trees$stump_diam[trees$status %in% c('Alive', 'Dead')] <- NA

ggplot(aes(x = stump_diam), data = subset(trees, !is.na(stump_diam))) +
  geom_histogram(binwidth = 5) +
  scale_y_log10() +
  scale_x_continuous(limits = c(0, 75))
```

Stump sizes exhibit a similar distribution.

```{r echo=FALSE}
ggplot(aes(x = user_type), data = trees) +
  geom_bar()
```

Here, we can see the number of trees cataloged by Parks Department staff,
census staff, and volunteers. Significant contributions were made by each of
the three groups. It might be interesting to see if volunteers' ratings of trees
differed at all from those of staff.

```{r echo=FALSE}
trees$problems[trees$status %in% c('Dead', 'Stump')] <- NA
common_problems <- names(tail(sort(table(trees$problems))))

ggplot(aes(x = problems),
       data = subset(trees, trees$problems %in% common_problems)) +
  geom_bar()
```

Looking at the six most common values in the 'problems' field, it appears that
the majority of trees had no reported problems, with 'Stones' and 'BranchLights'
being the most significant contributors otherwise. 'Stones' corresponds to the
presence of paving stones in the tree bed, which may cause root problems.
'BranchLights' refers to the presence of harmful light installations on the
trunk of the tree.

```{r echo=FALSE}
trees$spc_common[trees$status %in% c('Dead', 'Stump')] <- NA
common_trees <- names(tail(sort(table(trees$spc_common))))

ggplot(aes(x = spc_common),
       data = subset(trees, trees$spc_common %in% common_trees)) +
  geom_bar()
```

This barplot shows the counts of the six most common species of New York City
street trees. The most prevalent species are the London planetree, honey locust,
Callery pear, and pin oak; with over 50,000 specimens each, these four species
account for over 1/3 of the trees reported.

# Univariate Analysis

### What is the structure of your dataset?

The dataset contains information for 683,788 street trees in New York City.
Live and dead trees are included. 40 variables are included, mostly categorical
or location-related.

### What is/are the main feature(s) of interest in your dataset?

The most compelling part of this data is the potential to link the variables
analyzed above to location information - for example, species or health
distributions.

### What other features in the dataset do you think will help support your \
investigation into your feature(s) of interest?

The location data should be possible to combine with a city shapefile to
create interesting map plots.

### Did you create any new variables from existing variables in the dataset?

Not for univariate investigaion.

### Of the features you investigated, were there any unusual distributions? \
Did you perform any operations on the data to tidy, adjust, or change the form \
of the data? If so, why did you do this?

The only numerical variables were tree and stump diameter, which exhibited
relatively unsurprising distributions.

# Bivariate Plots Section

One of the questions that arose from the univariate exploration was whether
staff and volunteers exhibited differing behavior during evaluation of a tree's
status. This can be investigated using faceted bar plots.

```{r echo=FALSE, Bivariate_Plots}
ggplot(aes(x = health), data = subset(trees, status == 'Alive')) +
  facet_wrap(~ user_type, scales = 'free_y') +
  geom_bar()
```

The distributions seem quite similar - it looks like volunteers were equally
competent to staff in making assessments of tree health.

Let's start looking at the location data. The traditional way of creating map
plots in R is via the maps package. The package includes maps for countries and
states, but not cities, so for our purposes we need to find an appropriate
shapefile and load it (preferably into a dataframe). [This helpful post](https://github.com/tidyverse/ggplot2/wiki/plotting-polygon-shapefiles)
was consulted to learn how to do just that using the rgdal and maptools
packages. A shapefile for the five city boroughs can be obtained at the NYC
[Department of City Planning webpage](https://www1.nyc.gov/site/planning/data-maps/open-data/districts-download-metadata.page).

```{r echo=FALSE}
boroughs<-readOGR("Shapefiles/nybb.shp", layer="nybb")
boroughs@data$id = rownames(boroughs@data)
boroughs.points = fortify(boroughs, region="id")
boroughs.df = inner_join(boroughs.points, boroughs@data, by="id")
```

With the polygons laded, [this post](http://eriqande.github.io/rep-res-web/lectures/making-maps-with-R.html)
was used to better understand the plotting syntax for map shapes. Using
geom_polygon(), a simple map of New York City can be generated.

```{r echo=FALSE}
# geom_point(data = trees, aes(x = x_sp, y = y_sp), color = 'red')

nyc_base <- ggplot(data = boroughs.df,
                   aes(x = long, y = lat, group = group)) +
  geom_polygon(color = 'black', fill = 'gray') +
  coord_equal(ratio = 1) # fix map projection
nyc_base
```

By matching the borough names with those in the trees dataframe, we can use
map plots to visualize our variables with respect to location. For example,
the following plot illustrates the trees per area in each borough.

```{r echo=FALSE}
# Add this to a plot to drop axes/ticks but keep legend. From linked github post
ditch_the_axes <- theme(
  axis.text = element_blank(),
  axis.line = element_blank(),
  axis.ticks = element_blank(),
  panel.border = element_blank(),
  panel.grid = element_blank(),
  axis.title = element_blank()
  )

boroughs.df$numTrees <- as.numeric(table(trees$boroname)[boroughs.df$BoroName])

nyc_base + geom_polygon(data = boroughs.df,
                        aes(fill = numTrees / Shape_Area)) +
  scale_fill_gradient(low = "khaki1", high = "darkgreen") +
  ditch_the_axes
```

Judging by the legend, Manhattan has the greenest streets of the five boroughs,
with about 50% more trees per area than Staten Island. Given that this
dataset omits trees in parks and the like, it's a bit surprising that the most
densely populated borough also exhibits the highest street tree density.

Although borough visualizations are definitely of interest, the information they
contain could just as easily be gleaned from a table. It would be more
interesting to divide the map into further subsections and then explore further
variables. This can be done using a shapefile of the city's 59 community
districts.

Loading and preparing such a shapefile was completed in a similar fashion to
above, with the exception of obtaining the file from a [NYC OpenData map](https://data.cityofnewyork.us/City-Government/Community-Districts/yfnk-k7r4/data),
since the file from the City Planning website had some corrupt polygons.

```{r echo=FALSE}
communities<-readOGR("Shapefiles/nycd.shp",
                     layer="nycd")
communities@data$id = rownames(communities@data)
communities.points = fortify(communities, region="id")
communities.df = inner_join(communities.points, communities@data, by="id")
```

```{r echo=FALSE}
# geom_point(data = trees, aes(x = x_sp, y = y_sp), color = 'red')

nyc_cd_base <- ggplot(data = communities.df,
                      aes(x = long, y = lat, group = group)) +
  geom_polygon(color = 'black', fill = 'gray') +
  coord_equal(ratio = 1)
nyc_cd_base + ditch_the_axes
```

This map will allow for the display of more detailed information. Let's see
what the tree density looks like using this map.

```{r echo=FALSE}
communities.df$numTrees <- as.numeric(table(trees$cb_num)[as.character(communities.df$boro_cd)])

nyc_cd_base + geom_polygon(data = communities.df,
                           aes(fill = numTrees / shape_area)) +
  scale_fill_gradient(low = "khaki1", high = "darkgreen") +
  ditch_the_axes
```

This shows that the areas of greatest tree density tend to be towards the
center of the city, in areas such as Manhattan and much of Brooklyn.

```{r echo=FALSE}
volunteer_stats_cd <- trees %>%
  group_by(cb_num) %>%
  summarise(prop_volunteers = as.numeric(table(user_type)['Volunteer']) / n(),
            mean_size = mean(tree_dbh, na.rm = TRUE))

communities.df.volunteers <- inner_join(communities.df, volunteer_stats_cd,
                                        by = c('boro_cd' = 'cb_num'))
nyc_cd_base + 
  geom_polygon(data = communities.df.volunteers,
               aes(fill = prop_volunteers)) +
  scale_fill_gradient(low = 'blue', high = 'red') +
  ditch_the_axes
```

This plot illustrates the proportion of trees cataloged by volunteers (as
opposed to staff) in each community district. Volunteer activity seems to
mirror tree density, being higher in the central districts of the city.
The legend shows that these differences are quite large, with volunteers
cataloging around 10% of the trees in Staten Island as opposed to around 80% in
north Brooklyn. Volunteer engagement is much stronger in the areas
with higher tree density, even though tree density variation is not that large,
excepting central Manhattan. The next tree census may benefit from additional
promotion in the city's outer boroughs.

```{r echo=FALSE}
nyc_cd_base + 
  geom_polygon(data = communities.df.volunteers,
               aes(fill = mean_size)) +
  scale_fill_gradient(low = 'blue', high = 'red') +
  ditch_the_axes
```

Here, the average tree diameter is displayed for each community district. The
average street tree in much of Queens and Brooklyn is over a foot in diameter,
while Manhattan and West Bronx street trees average significantly smaller; this
makes sense given the densely populated developments of those neighborhoods.

```{r echo=FALSE}
diversity_by_cd <- trees %>%
  group_by(cb_num) %>%
  summarise(num_species = length(table(spc_common)[table(spc_common) > 0]))

communities.df.diversity <- inner_join(communities.df, diversity_by_cd,
                                       by = c('boro_cd' = 'cb_num'))
nyc_cd_base + 
  geom_polygon(data = communities.df.diversity,
               aes(fill = num_species)) +
  scale_fill_gradient(low = 'tomato4', high = 'chartreuse3') +
  ditch_the_axes
```

This map illustrates street tree diversity as the number of species in each
district. From this visualization, it is apparent that Manhattan and Bronx
districts contain fewer species of street tree on average. However, given that
the community districts vary significantly in area, it might also be useful
to plot this information per unit area by dividing num_species by each shape's
area.

```{r echo=FALSE}
nyc_cd_base + 
  geom_polygon(data = communities.df.diversity,
               aes(fill = num_species/shape_area)) +
  scale_fill_gradient(low = 'tomato4', high = 'chartreuse3') +
  ditch_the_axes
```

This plot tells a different story - Manhattan and West Bronx are among the
most diverse areas, judging by species per area. This may correspond to the
earlier finding that Manhattan had the most trees per area while Staten Island
had the least - more plantings is equated with greater diversity.

Moving on from diversity, are there any other interesting patterns at the
species level? One idea might be to plot a map with the most common tree
species in each district.

```{r echo=FALSE}
# Base r does not contain a mode function. This returns a single mode
# From https://stackoverflow.com/questions/2547402/is-there-a-built-in-function-for-finding-the-mode
Mode <- function(x) {
  ux <- unique(x)
  ux[which.max(tabulate(match(x, ux)))]
}

species_by_cd <- trees %>%
  group_by(cb_num) %>%
  summarise(abundant_tree = Mode(spc_common))

communities.df.species <- inner_join(communities.df, species_by_cd,
                                     by = c('boro_cd' = 'cb_num'))
nyc_cd_base + 
  geom_polygon(data = communities.df.species,
               aes(fill = abundant_tree)) +
  ditch_the_axes
```

This plot displays the most abundant tree species in each district. The five
species represented are the five most common overall, and the city-wide counts
are as follows:

```{r echo=FALSE}
head(sort(table(trees$spc_common), decreasing = TRUE), 5)
```

Given these numbers, the species distributions are quite interesting. Manhattan,
West Bronx, and North Brooklyn are dominated by honey locusts, while the
surrounding areas are predominantly planted with London planetrees. This
seems to mirror the average trunk size plot created earlier. Is the planting
distribution related to tree size?

```{r}
summary(trees$tree_dbh[trees$spc_common == 'honeylocust'])
```

```{r}
summary(trees$tree_dbh[trees$spc_common == 'London planetree'])
```

Honey locust trees are about half as large (judging by diameter) as london
planetrees, on average. This could mean that honey locusts are planted more
often in the aforementioned congested parts of the city, while planetrees are
better suited to more open neighborhoods.

To better compare the distributions of the two species, we can
make a map plot which only considers planetrees and honey locusts.

```{r echo=FALSE}
dddd <- subset(trees, status = 'Alive') %>%
  group_by(cb_num) %>%
  summarise(pct_pt = as.numeric(table(spc_common)['London planetree']/(table(spc_common)['honeylocust'] +  table(spc_common)['London planetree'])))

communities.df.dddd <- inner_join(communities.df, dddd,
                                  by = c('boro_cd' = 'cb_num'))
nyc_cd_base + 
  geom_polygon(data = communities.df.dddd,
               aes(fill = pct_pt)) +
  scale_fill_gradient(low = 'yellow', high = 'blue') +
  ditch_the_axes
```

This plot shows the proportion of planetrees relative to honey locusts in each
district. In much Manhattan, there are at least 4 honey locusts to every
planetree (judging by the proportion of 0.2 on the legend). The ratio is closer
in much of the rest of the city, with only a few districts colored blue or
deep purple (indicating significant prevalence of planetrees).

To verify the assumptions I've been making about congestion in Manhattan and
West Bronx, it might be a good idea to try to factor in some population data
after I've had enough fun with these exploratory map plots.

Lastly, it would be interesting to revisit tree health in the context of
location within the city.

```{r echo=FALSE}
health_by_cd <- subset(trees, status = 'Alive') %>%
  group_by(cb_num) %>%
  summarise(pct_healthy = as.numeric(table(health)['Good']) / n())

communities.df.health <- inner_join(communities.df, health_by_cd,
                                     by = c('boro_cd' = 'cb_num'))
nyc_cd_base + 
  geom_polygon(data = communities.df.health,
               aes(fill = pct_healthy)) +
  scale_fill_gradient(low = 'orangered', high = 'dodgerblue') +
  ditch_the_axes
```

This plot illustrates the proportion of healthy (health = 'Good') trees in each
district. Much of Manhattan and West Bronx contains disproportionately unhealthy
trees. Most notably, the bottom-right district - Rockaway, Queens - has a
surprisingly low 'health rating'. Only about 60% of trees were judged as being
in good health. Why might this be the case? Bringing some more variables into
the equation may help answer this question.

# Bivariate Analysis

### Talk about some of the relationships you observed in this part of the \
investigation. How did the feature(s) of interest vary with other features in \
the dataset?

Many variables exhibited interesting trends when related to location. Volunteer
activity was much higher in the central parts of the city. Variables such as
species diversity, tree size, and tree health seemed to have similar
distributions with respect to location.

### Did you observe any interesting relationships between the other features \
(not the main feature(s) of interest)?

The Rockaway peninsula in Queens had a very large percentage of unhealthy trees.
This would be an interesting phenomena to explore further.

### What was the strongest relationship you found?

The relationship between location and volunteer activity was quite profound,
varying from less than 20% to over 80% depending on the community district.

# Multivariate Plots Section

The bivariate exploration raised two location-related questions that will be
further explored in this section:

1) Why do inner-city districts in Manhattan, West Bronx, etc. exhibit different
species distributions and tree health?

2) Why are trees in Rockaway, Queens faring so poorly?

Starting with the first question, our intuition was that street trees fared
worse in more congested districts. To justify this intuition, we can use
another dataset provided by the city's OpenData initiative that includes
populations for each of the 59 community districts, recorded once every decade
(most recently in 2010). This data can then be joined to the communities
shapefile and overlaid on the tree health plot using geom_point.

```{r echo=FALSE, Multivariate_Plots}
population <- read.csv("Data/New_York_City_Population_By_Community_Districts.csv")
str(population)
```

```{r echo=FALSE}
boro_to_id <- function(df, boroname, cdnum) {
  boro_additives <- c(100, 200, 300, 400, 500)
  names(boro_additives) <- c("Manhattan",
                            "Bronx",
                            "Brooklyn",
                            "Queens",
                            "Staten Island")
  boronum <- as.numeric(boro_additives[as.character(df[[boroname]])])
  df[["borocode"]] <- boronum + df[[cdnum]] 
  df
}

population <- boro_to_id(population, "Borough", "CD.Number")
```

```{r echo=FALSE}
communities.df.pop <- inner_join(communities.df, population,
                                 by = c("boro_cd" = "borocode"))
community.centers <- communities.df.pop %>%
  group_by(boro_cd) %>%
  summarise(x = mean(range(long)),
            y = mean(range(lat)))

community.centers <- as.data.frame(community.centers)
communities.df.pop <- inner_join(communities.df.pop, community.centers,
                                 by = "boro_cd")
communities.df.pop$pop_density <- communities.df.pop$X2010.Population / communities.df.pop$shape_area

ggplot(data = communities.df, aes(x = long, y = lat, group = group)) +
  geom_polygon(color = 'black', fill = 'gray') +
  geom_polygon(data = communities.df.health,
               aes(fill = pct_healthy)) +
  scale_fill_gradient(low = 'orangered', high = 'dodgerblue') +
  geom_point(data = communities.df.pop,
             aes(x = x,
                 y = y,
                 size = pop_density,
                 group = NULL)) +
  coord_equal(ratio = 1) +
  ditch_the_axes
```

Judging by this plot, it seems that the theory of population density affecting
tree health holds true in Manhattan and West Bronx. In those districts,
the population densities are indeed very high compared to much of the rest of
the city.

However, the
relationship appears to be be weaker in Brooklyn and Queens - there are subtle
trends in health that do not seem to directly related to population. The most
striking example is the aforementioned Rockaway Peninsula - population density
is quite low, yet the street tree health is worse on average than in any other
district.

We can further investigate this district by subsetting the data and comparing
tree health to more variables. Maybe some species tend to fare worse than
others. A species-specific blight or insect infestation of an abundant species
could bring down the average significantly.

```{r}
# (confirmed by plotting district 414)
ra_trees <- subset(trees, cb_num == 414 & status == 'Alive')
summary(table(ra_trees$spc_common) > 0)
```

Rockaway is actually one of the most diverse districts, containing 109 of the
total 133 species of NYC street tree. Let's plot the average health for the
most common species (over 100 individuals in the district).

```{r echo=FALSE}
tree_counts <- table(ra_trees$spc_common)
common_trees <- unique(ra_trees$spc_common[tree_counts[ra_trees$spc_common] > 100])
rockaway_commons <- subset(ra_trees, spc_common %in% common_trees)

common_health <- rockaway_commons %>%
  group_by(spc_common) %>%
  summarise(mean_health = as.numeric(table(health)['Good']) / n(),
            n = n())

ggplot(data = common_health,
       aes(x = reorder(spc_common, mean_health), y = mean_health)) +
  geom_point(aes(size = n)) +
  geom_hline(yintercept = table(trees$health)['Good'] / nrow(trees),
             color = "blue") +
  coord_flip()
```

This plot illustrates the average health of the most common species, with the
dot size corresponding to tree count and the blue line representing the
citywide average. (Again, health is measured as the number of live trees rated
as being in 'Good' health divided by the total number of live trees.) Clearly,
there is quite a bit of variation in health by species, but only two of the
24 species with counts over 100 exhibit health proportions greater than the
citywide average. Notably, the two most common species in the city both
fare poorly in Rockaway: about 50% of London planetrees and below 70% of
honey locusts are in good health.

This plot indicates that many species of street tree suffer in this particular
district, rather than one or two abundant species bringing down the average as
theorized. An alternate theory is that the problems recorded for each tree,
such as stones in planting beds and lighting affixed to the tree, occur in
higher proportions in Rockaway. This can be investigated using another simple
map plot.

```{r echo=FALSE}
problems_by_cd <- subset(trees, status = 'Alive') %>%
  group_by(cb_num) %>%
  summarise(pct_problematic = 1 - (as.numeric(table(problems)['None']) / n()))

communities.df.problems <- inner_join(communities.df, problems_by_cd,
                                     by = c('boro_cd' = 'cb_num'))
nyc_cd_base + 
  geom_polygon(data = communities.df.problems,
               aes(fill = pct_problematic)) +
  scale_fill_gradient(low = 'dodgerblue', high = 'orangered') +
  ditch_the_axes
```

Actually, Rockaway has a percentage of problematic trees on the lower end of the
scale, so this theory doesn't hold up either.

The best remaining theory seems to be that Rockaway is simply an inhospitable
environment for street trees. Perhaps storms are more likely to adversely
affect the exposed trees in this beach-lined district. Another possibility is
that young street trees are not as well cared for as described on the
[Parks Department webpage](https://www.nycgovparks.org/trees/street-tree-planting)
during their first couple of years, a responsibility belonging to the street
tree contractors.

# Multivariate Analysis

### Talk about some of the relationships you observed in this part of the \
investigation. Were there features that strengthened each other in terms of \
looking at your feature(s) of interest?

The addition of the population data seemed to support the relationship between
population density and tree health for Manhattan and Bronx, but the
relationship was not as strong in other boroughs, the most notable outlier
being the Rockaway Peninsula at the south of Queens. In Rockaway, poor tree
health was found to be representative of many of the common species, leading
to theories outside the scope of this data.

------

# Final Plots and Summary

### Plot One
```{r echo=FALSE, Plot_One}
final1 <- nyc_cd_base + 
  geom_polygon(data = communities.df.volunteers,
               aes(fill = prop_volunteers)) +
  scale_fill_gradient(low = 'blue', high = 'red') +
  labs(title = "Volunteer Activity in 2015 NYC Street Tree Count") +
  theme(plot.title = element_text(hjust = 0.5)) +
  ditch_the_axes
final1$labels$fill <- "Proportion of\nvolunteers"
final1
```

### Description One

This plot illustrates the proportion of trees counted by volunteers in each
district during the 2015 street tree census. Volunteer activity varied
significantly: central districts were catalogued almost entirely by
volunteers, while Staten Island and outer Queens and Bronx were largely covered
by staff. This plot could be used to determine areas in which volunteer
engagement can be improved.

### Plot Two
```{r echo=FALSE, Plot_Two}
final2 <- ggplot(data = communities.df, aes(x = long, y = lat, group = group)) +
  geom_polygon(color = 'black', fill = 'gray') +
  geom_polygon(data = communities.df.health,
               aes(fill = pct_healthy)) +
  scale_fill_gradient(low = 'orangered', high = 'dodgerblue') +
  geom_point(data = communities.df.pop,
             aes(x = x,
                 y = y,
                 size = pop_density,
                 group = NULL)) +
  coord_equal(ratio = 1) +
  labs(title = "Street Tree Health and Population Density in Community Districts") +
  theme(plot.title = element_text(hjust = 0.5)) +
  ditch_the_axes
final2$labels$fill <- "Proportion of\ntrees in\ngood health"
final2$labels$size <- "Relative\npopulation\ndensity"
final2
```

### Description Two

This plot shows the proportion of trees in good health in each district,
overlaid with dots whose size corresponds to the population density of each
district. The relatively poor tree health in Manhattan and West Bronx seems to
relate to the very high population densities of those neighborhoods. However,
it appears to be possible to sustain healthy street tree populations even with
high population density, as evidenced by the very healthy districts in central
Brooklyn with moderately high population density. Lastly, the southeastern
Rockaway district is a notable outlier.

### Plot Three
```{r echo=FALSE, Plot_Three}
final3 <- ggplot(data = common_health,
       aes(x = reorder(spc_common, mean_health), y = mean_health)) +
  geom_point(aes(size = n)) +
  geom_hline(yintercept = table(trees$health)['Good'] / nrow(trees),
             color = "blue") +
  labs(x = "Species",
       y = "Proportion in good health",
       title = "Street Tree Health by Species in Rockaway, Queens") +
  coord_flip()
final3$labels$size <- "Count"
final3
```

### Description Three

This plot illustrates the proportion of trees in good health for the most
common species in the Rockaway district, with dot sizes corresponding to tree
count and the citywide proportion plotted as the blue line. This plot was
made to address the possibility that poor health in one or a few abundant
species brought down the district's tree health rating. As it turns out, the
overwhelming majority of the district's primary tree species fell below the
citywide health rating, indicating that the problems with tree health in
Rockaway are unlikely to be species-related.

------

# Reflection

The most exciting parts of this investigation were learning to use map plots
for compelling visualizations and leveraging data from two different datasets
to make a stronger argument in the multivariate analysis. The function to
generate the community district code for the population data was surprisingly
difficult until I came across [a Stack Overflow post](https://stackoverflow.com/questions/2641653/pass-a-data-frame-column-name-to-a-function)
discussing how to pass dataframe column names to functions. I was interested
in adding more socioeconomic data to investigate tree health in different
districts, but wasn't able to find an appropriate dataset - future work might
involve scraping additional information and adding it to map plots.